Scalar and Representative Models for Polarimetric SAR Data

### 2013-07-24 paper_dev/article_pp_ivm.tex
updated for Ian to change

%ifdef FINAL
\documentclass[journal]{IEEEtran}
%endif
%ifdef REVIEW
\documentclass[journal,12pt,draftcls,onecolumn]{IEEEtran}
%endif
%\documentclass[a4paper, 10pt, conference]{ieeeconf}

@@

%opening
\title{
  Scalar and Representative Observables for Polarimetric SAR Data
}

%\author{Thanh-Hai Le, Ian McLoughlin}
\author{Thanh-Hai~Le,
        Ian~McLoughlin, 
	and Chan-Hua~Vun%
\thanks{Thanh-Hai~Le and Chan-Hua~Vun are with School of Computer Engineering, 
Nanyang Technological University, Singapore. Ian~McLoughlin is with School of Information Science and Technology,
University of Science and Technology of China.
}% <-this % stops a space
%\thanks{The authors wish to thank Dr. Ken-Yoong Lee and Dr. Timo Brestchneider of EADS Innovation-Works Singapore for 
%	providing us the RADAR-SAT2 imagery used in this paper. }% <-this % stops a space
\thanks{Manuscript received ?, 2013; revised ?.}}

\markboth{IEEE Journal on Selected Topic in Applied Earth Observation and Remote Sensing,~Vol.~?, No.~?, ?~2013}%
{ Le \MakeLowercase{\textit{et al.}}:  Scalar and Representative Observables for Polarimetric SAR Data}

@@

\begin{abstract}
This paper presents several novel scalar statistical models to represent the determinant of both partial and full polarimetric SAR (POLSAR) covariance matrices.
Compared to other scalar statistical models for POLSAR,
  the proposed models are highly representative of the multi-dimensional data
  and enable useful discrimination measures to be easily determined. %consequently they lead to useful discrimination measures.
The representative power of these models is explained  
when these multi-dimensional models are collapsed to single dimensional versions,
  they are shown to exactly match the traditional statistical models for the representative observable of SAR intensity.
The paper also validates the theoretical models together with their derived discrimination measures using real world POLSAR data.  
%IVMThe proposed generic models are envisioned to enable the adaptation of many existing SAR data processing techniques for use with POLSAR data.
One major benefit of this paper is that these proposed generic models can facilitate the adaptation of existing SAR techniques to higher-dimensional POLSAR data.
To illustrate, an example is given where the determinant-ratio and its additive version, contrast, are applied in the evaluation of POLSAR speckle filters.
\end{abstract}

@@

\section{Introduction}

During the past decades, exponential growth in computing power has allowed once computationally-demanding Synthetic Aperture Radar (SAR)
technology to become a feasible and preferred technique for earth observation applications.
SAR technology has since extended in a few directions, one of which is polarimetric SAR (POLSAR).
POLSAR extends SAR by exploiting the natural polarization property of Electro-Magnetic (EM) waves,
  leading to the availability of multi-channel POLSAR data, compared to traditional one-channel SAR data.

While statistical models are important in understanding the stochastic nature of both SAR and POLSAR, in extending our understanding towards multi-dimensional POLSAR, an important issue needs to be addressed.
That is there now exists not one but many observable quantities in multi-channel POLSAR (as opposed to single-channel SAR intensity).
The research community has responded largely by developing different statistical models for various POLSAR observables.
However for a statistical model to be useful, scalar discrimination measures generally need to be derivable from it.
Thus, practical applications for POLSAR data processing require dis-similarity measures that are scalar, consistent and preferably homoskedastic on the one hand.
On the other hand, the observable quantity being modelled should be naturally representative of the high dimensionality POLSAR data.

This paper presents novel statistical models for multi-channel POLSAR data as generalisations of the commonly used and representative statistical models for the one-channel SAR data.
This approach makes feasible the adaptation of many existing SAR processing techniques to be applied to POLSAR data.
After section \ref{sec:lit_review} reviews related work in published literature, statistical models for the heteroskedastic POLSAR determinant and its homoskedastic log-domain equivalent, together with several other discrimination measures are derived in section  \ref{sec:theoretical_model}.
The representative properties of these observables are then demonstrated in section \ref{sec:sar_special_case_of_polsar}
  where the multi-dimensional POLSAR data is collapsed to a one-dimensional SAR scenario.
With this transformation, the determinant of the POLSAR covariance matrix is shown to be equivalent to SAR intensity,
%  which also brings many standard statistical models used with SAR  under the umbrella of these generic POLAR models.
  and allow many conventional SAR statistical models to stay under the umbrella of these generic POLAR models.
The proposed theoretical models are then validated against real life practical data in section \ref{sec:polsar_models_validation}.
Section \ref{sec:discussion} briefly describes how evaluation of an existing SAR data processing technique, the speckle filter, can be extended to  POLSAR data processing using this approach.
This is followed by a conclusion in Section \ref{sec:conclusion}.

@@

This section outlines related published work.
In particular, section \ref{sec:lit_models} reviews various scalar statistical models used for different POLSAR observables.
Its purpose is to show that none of those  models have led to statistically consistent discrimination measures. 
Section \ref{sec:lit_measures} further strengthens those findings by discussing discrimination measures that have been proposed for POLSAR,
  and demonstrates that almost all of them are based on the likelihood statistical test for complex Wishart distribution.
However, while an exact statistical distribution is needed for the test,
  only asymptotic distribution is used in the original work \cite{Conradsen_2003_TGRS_4}.

\subsection{Scalar Observables for POLSAR Data and their Statistical Models}
\label{sec:lit_models}

Different target decomposition theorems have identified many possible scalar observables for complex POLSAR data.
In \cite{Alberga_2008_IJRS_4129}, the performance of different scalar POLSAR observables is evaluated for classification purposes.
While many scalar observables for POLSAR were presented, their corresponding statistical models and classifiers were not available.
Furthermore, at its conclusion, the paper indicated that it is impossible to identify a single best representation.
Although, to be fair, those observables were identified to describe a decomposed portion of the complex POLSAR data,
  rather than providing a single unified representation of the POLSAR data.

@@

More recently, statistical models for
  each element of the POLSAR covariance matrix $S_{pq}S_{rs}^*$ \cite{Lopez-Martinez_2003_TGRS_2232}
  as well as for the largest eigen-value of the covariance matrix $\lambda_1$ \cite{Erten_2012_Sensors_2766} have been proposed.
While these models undoubtedly help in understanding POLSAR data,
  none of these observables have been shown to meet the dual criteria of
  1) resulting in statistically consistent discrimination measures and
  2) being representative of the complex POLSAR data.

@@

The commonly used measure of distance for matrices are either the Euclidean or Manhattan distances, defined as:
\begin{align}
  d(C_x,C_y) &= \sum_{i,j} |\mathbb{R} (C_x - C_y)_{i,j}| + \sum_{i,j} |\mathbb{I} (C_x - C_y)_{i,j}| \\
  d(C_x,C_y) &= \sqrt{\sum_{i,j} |C_x - C_y|_{i,j}^2 }
\end{align}
where $C_{i,j}$ denotes the (i,j) elements of the POLSAR covariance matrix C,
 $||$ denotes absolute values
and $\mathbb{R},\mathbb{I}$ denote the real and imaginary parts respectively.
However, in the context of POLSAR% covariance matrix
, these dis-similarity measures are not widely used 
  mainly due to the multiplicative nature of the noisy data.

In the field of POLSAR, the Wishart distance is probably the most widely used, as part of the well-known Wishart classifier \cite{Lee_1999_TGRS}.
It is defined \cite{Lee_1994_IJRS_2299} as:
\begin{equation}
  d(C_x,C_y) = \ln|C_y| + tr(C_xC_y^{-1})
\end{equation},
where $tr(C)$ denotes the trace of the matrix C. 
As a measure of distance, its main disadvantage is that $d(C_y,C_y) = \ln|C_y| \neq 0$.

Recent works have suggested alternative dissimilarity measures including the symmetric and asymmetric refined Wishart distance \cite{Anfinsen_2007_ESA_POLINSAR},
\begin{align}
  d(C_x,C_y) &= \frac{1}{2} tr(C_x^{-1}C_y + C_y^{-1}C_x) - d \\
    d(C_x,C_y) &= \ln|C_x| - \ln|C_y| + tr(C_xC_y^{-1}) - d
\end{align}
the Bartlett distance \cite{Kersten_2005_TGRS_519},
  \begin{align}
  d(C_x,C_y) &= 2 \ln |C_{x+y}| - \ln |C_x| - \ln |C_y| - 2d\ln2
  \end{align}
the Bhattacharyya distance \cite{Lee_2011_IGARSS_3740},
\begin{equation}
  r(C_x,C_y) = \frac{|C_x|^{1/2} |C_y|^{1/2}}{|(C_x+C_y)/2|}
\end{equation}
and the Wishart Statistical test distance \cite{Cao_2007_TGRS_3454}
\begin{equation}
  d(C_x,C_y) = (L_x + L_y) \ln|C| - L_x \ln|C_x| - L_y\ln|C_y|
\end{equation}
.

@@

Closer investigation of these dis-similarity measures reveals that most are related in some ways. %to each other.
The Bhattacharyya distance is easily shown to be related to the Barlett distance.
At the same time the Barlett distance can be considered a special case of the Wishart Statistical Test distance,
  when the two data sets have the same number of looks, i.e. $L_x=L_y$.
The close relation among the measures is further supported by the fact that
  all of their proposing papers refered the same statistical model developed in \cite{Conradsen_2003_TGRS_4} as a foundation.
In \cite{Conradsen_2003_TGRS_4}, to determine if the two scaled multi-look POLSAR covariance matrix $Z_x$ and $Z_y$,
  which have $L_x$ and $L_y$ as the corresponding number of looks,
  come from the same underlying stochastic process,
the likelihood ratio statistics for POLSAR covariance matrix is considered:  
\begin{equation*}
  Q = \frac{(L_x+L_y)^{d \cdot (L_x+L_y)}}{L_x^{d \cdot L_x} L_y^{d \cdot L_y}} \frac{|Z_x|^{L_x} |Z_y|^{L_y} }{|Z_x+Z_y|^{(L_x+L_y)}}
\end{equation*}

Taking the log-transformation of the above equation, and note that $C_{vx} = Z_x / L_x$, $C_{vy} = Z_y / L_y$ and $C_{vxy} = (Z_x + Z_y)/(L_x + L_y)$ then:
\begin{eqnarray}
  Q &=& \frac{|C_{vx}|^{L_x} \cdot |C_{vy}|^{L_y} }{|C_{vxy}|^{L_x + L_y}} \label{eqn:ori_likelyhood_stats} \\
  \ln Q &=& L_x \ln |C_{vx}| + L_y \ln |C_{vy}| - (L_x + L_y) \ln |C_{vxy}| \label{eqn:log_likelyhood_stats}
\end{eqnarray}

To detect changes, a test statistic is developed based on this measure of distance.
This means a distribution is to be derived for the dissimilarity measure.
However, in the original work \cite{Conradsen_2003_TGRS_4}, only an asymptotic distribution has been proposed.

@@

In this section, several scalar statistical models for POLSAR are presented.
Section \ref{sec:model_basic} briefly introduces the basic foundations, concepts and notation used in this paper, %reviews related work in published literature,
  while section \ref{sec:polsar_heterosked_model_and_log_transform}
  presents the theoretical models together with mathematical evidence for two related points.
The first point is that the POLSAR data is multiplicative and heteroskedastic in its original domain.
And the second conclusion states that log-transformation converts it into an additive and homoskedastic model.
Some consistent measures of distance are then derived and presented in section \ref{sec:distance_measure}.

@@

The sample POLSAR covariance matrix is formed as the mean of Hermitian outer product of independent single-look scattering vectors,
\begin{equation}
  C_v = \langle ss^{*T} \rangle = \frac{1}{L} \sum^L_{i=1}s_is_i^{*T}
\end{equation}
where $s_i$ denotes the single-look scattering vector,
  which equals $s_{part}$ for the case of partial POLSAR or
  $s_{full}$ for the case of full polarimetry,
and $L$ is the number of looks.

@@

\subsection{Difference between the theoretical assumptions and the conditions found in practice}  
\label{sec:improve_the_match_bw_theory_practice}

>>>>

%\documentclass[a4paper,10pt]{article}
%ifdef FINAL
\documentclass[journal]{IEEEtran}
%endif
%ifdef REVIEW
\documentclass[journal,12pt,draftcls,onecolumn]{IEEEtran}
%endif
%\documentclass[a4paper, 10pt, conference]{ieeeconf}

@@

%\nonstopmode

%opening
\title{
  %Consistent and Homoskedastic Measures of Distance for Heteroskedastic POLSAR Data
  Scalar and Representative Models for Polarimetric SAR Data
}

%\author{Thanh-Hai Le, Ian McLoughlin}
\author{Thanh-Hai~Le,
        Ian~McLoughlin, 
	and Chan-Hua~Vun%
\thanks{Thanh-Hai~Le and Chan-Hua~Vun are with School of Computer Engineering, 
Nanyang Technological University, Singapore. Ian~McLoughlin is with School of Information Science and Technology,
University of Science and Technology of China.
}% <-this % stops a space
%\thanks{The authors wish to thank Dr. Ken-Yoong Lee and Dr. Timo Brestchneider of EADS Innovation-Works Singapore for 
%	their initial discussions and for providing us the RADAR-SAT2 imagery used in this paper. }% <-this % stops a space
\thanks{Manuscript received ?, 2013; revised ?.}}

\markboth{Transactions on Geoscience \& Remote Sensing,~Vol.~?, No.~?, ?~2013}%
{ Le \MakeLowercase{\textit{et al.}}:  Scalar and Representative Models for Polarimetric SAR Data}

@@

\begin{abstract}
This paper presents several novel scalar statistical models to represent the determinant of both partial and full polarimetric SAR (POLSAR) covariance matrices.
Compared to other scalar statistical models for POLSAR,
  the proposed models are highly representative of the multi-dimensional data
  and enable useful discrimination measures to be easily determined. %consequently they lead to useful discrimination measures.
When the multi-dimensional models are collapsed to a single dimension,
  they are shown to exactly match the traditional statistical models for SAR intensity.
%The representative power of these multi-dimensional models is illustrated
%  when they are collapsed into the special case of one-dimensional,
%  the special-case statistical models then match perfectly with the traditionally used statistical models for SAR intensity.
%These theoretical models, together with their derived discrimination measures, are then validated against practically captured POLSAR data.
The paper validates the theoretical models together with their derived discrimination measures using real world POLSAR data.  
%IVMThe proposed generic models are envisioned to enable the adaptation of many existing SAR data processing techniques for use with POLSAR data.
The intention is that these proposed generic models can facilitate the adaptation of existing SAR techniques to higher-dimensional POLSAR data.
To illstrate, an example is given where the determinant-ratio and its additive version, contrast, are applied in the evaluation of POLSAR speckle filters.
%For example, similar to the way that the intensity ratio is routinely used to evaluate SAR speckle filters,
%  this paper describes how the determinant-ratio and its additive version, contrast, can be used to evaluate POLSAR speckle filters.
%  While statistical models are important in understanding the stochastic nature of Polarimetric SAR (POLSAR),
%  scalar discrimination measures are routinely used in processing this multi-dimensional data.
%  Ideally speaking, the latter should be drived from the former.
%  As far as we know, however, it has not exactly been so for the case of POLSAR.
%  In this paper, a couple of statistical models for the determinant of the POLSAR covariance matrix and its log-transformed derivations are proposed.
%  These scalar models for the determinant of the POLSAR covariance matrix are shown to be respresentative
%    since this observable is neatly transformed into the traditional SAR intensity
%    when the generic multi-dimensional polarimetry is collapsed into the case of one-dimensional SAR. %case of traditional SAR data.
%  Similar to SAR, the generic model for POLSAR is also shown to be multiplicative and heteroskedastic in the original domain 
%    and the logarithmic transformation converts this into a more familiar additve and homoskedastic model.
%  Moreover, the scalar property of these statistical models lead to the derivation of several statistical discrimination measures. %homoskedastic measures of distance.
%  All of these statistical models are validated on real-life captured data
%    and they are shown to be applicable even when the theoretical ``independent samples'' assumption is violated in practice.
%  The combination of scalar, additive and homoskedastic properties of the derived statistical models bring about a couple of benefits, which are also described in this paper.
%  First, compared to the original multiplicative domain, %dis-criminative measure of ratio,
%    the additive log-transformed fit more naturally with the linear nature of digital imagery.
%  Second, since the Gauss-Markov theorem is again applicable in the homoskedastic log-transformed domain,
%    the use of Mean Squared Error (MSE) as an reliable objective function for POLSAR speckle filters is briefly explored.
\end{abstract}

@@

\section{Introduction}
%\section{POLSAR data: statistical analysis and logarithmic transformation}

During the past decades, exponential growth in computing power has allowed once computationally-demanding Synthetic Aperture Radar (SAR)
technology to become a feasible and preferred technique for earth observation applications.
SAR technology has since extended in a few directions, one of which is polarimetric SAR (POLSAR).
POLSAR incorporates multiple channels to represent the natural polarization properties of reflected and sensed Electro-Magnetic (EM) waves, unlike the single dimension of SAR intensity.

%In the past decades, the exponential growth in computational power has made the once overly-excessive computationaly-demanding SAR technology now become a feasible and preferred earth observation solution.
%As state-of-the-art technology, the SAR technique has been extended in a few directions, one of which is the polarimetric SAR.
%POLSAR is the natural extension from SAR exploiting the natural polarization property of Electro-Magnetic (EM) waves.
%Similar to the extension from black-and-white images to color photography, the polarimetric extension brought about the multi-channels POLSAR data as compared to the traditional one-channel SAR data.

%With (POL)SAR data becomes cheaper and more available, the recent research emphasis is on understanding and developing applications for the data. %it is of great important to understand the data.
%The complex data, however is stochastic by nature, due to the interference phenomena of EM waves. %, SAR measurements are stochastic by nature.
%Under this condition, it is very important to understand the statistical models for (POL)SAR data.

%While the statistical models for homogeneous areas of one-dimensional SAR data have been developed,
%  in extending the model toward heterogeneous images, the impact of heteroskedasticity has virtually been ignored.
%Specifically the concept of distance as a measure of dis-similarity is seriously flawed, as it is widely known that ratio offers a better discrimination measure TODOCITE.
%This thesis starts off by proposing the logarithmic transformation which is shown converting the SAR data into an homoskedastic model.
%From this log-transformed model, a few consistent measures of distance is proposed. 
%in a sense the distance concept is re-emerged, as the log-transformation converts the widely used ratio into standard subtraction, and thus distance.
While statistical models are important in understanding the stochastic nature of both SAR and POLSAR, in extending our understanding towards multi-dimensional POLSAR, an important issue needs to be addressed.
Namely that there now exists not one but many observable quantities in multi-channel POLSAR (as opposed to single-channel SAR intensity).
The research community has responded largely by developing different statistical models for various POLSAR observables.
However for a statistical model to be useful, scalar discrimination measures generally need to be derivable from it.
Thus, practical applications for POLSAR data processing require dis-similarity measures that are scalar, consistent and preferably homoskedastic on the one hand.
On the other hand, the observable quantity being modelled should be naturally representative of the high dimensionality POLSAR data.

%While statistical models are important in understanding the stochastic nature of both SAR and POLSAR,
%  in extending our understanding towards multi-dimensional POLSAR data, an important issue needs to be addressed.
%The trouble that the high dimensional data bring about is that there exists not one, as the intensity in the one-channel SAR, but many observables quantities in multi-channel POLSAR.
%While different statistical models have been developed for different POLSAR observables,
%  for a statistical model to be useful, scalar discrimination measures need to be derived from it.
%Thus practical application for POLSAR data processing requires the dis-similarity measure to be scalar, consistent and preferably homoskedastic on one hand.
%On the other hand, the observable quantity being modelled needs to be naturally representative for the high dimensional POLSAR data.

%In this paper, the determinant of the POLSAR covariance matrix is proposed as an observable quantity to study.
%Statistical models for this heteroskedastic POLSAR determinant and homoskedastic log-determinant, together with several discrimination measures, are then derived in section \ref{sec:theoretical_model}.
%The representative power of this observable is demonstrated in section \ref{sec:sar_special_case_of_polsar},
%  where the multi-dimensional POLSAR data is collapsed into the traditional one-dimensional SAR scenario.
%Under this transformation, the determinant of the POLSAR covariance matrix is converted into the standard SAR intensity
%Section \ref{sec:sar_special_case_of_polsar} describes how
%  which puts the standard statistical models for SAR to be within the natural coverage of this generic models for POLSAR.
%All of these theoretical developments are then validated against real-life practical data in section  \ref{sec:polsar_models_validation}.
%Finally section \ref{sec:discussion_conclusion} concludes the paper by briefly describing how an existing SAR data processing technique can be adapted for POLSAR data.

This paper presents novel statistical models for multi-dimensional POLSAR data as generalisations of the commonly used one-dimensional SAR statistical model.
This approach makes feasible the adaptation of many existing SAR processing techniques, to be applied to POLSAR data.
Specifically, statistical models for the heteroskedastic POLSAR determinant and its homoskedastic log-domain equivalent, together with several other discrimination measures are derived in section  \ref{sec:theoretical_model}.
The representative properties of these observables are then demonstrated in section \ref{sec:sar_special_case_of_polsar}
  where the multi-dimensional POLSAR data is collapsed to a one-dimensional SAR scenario.
With this transformation, the determinant of the POLSAR covariance matrix is shown to be equivalent to SAR intensity,
  which also brings many standard statistical models used with SAR  under the umbrella of these generic POLAR models.
The proposed theoretical models are then validated against real life practical data in section \ref{sec:polsar_models_validation}.
Section \ref{sec:discussion} briefly describes how evaluation of an existing SAR data processing technique, the speckle filter, can be extended to  POLSAR data processing using this approach.
This is followed by a conclusion in Section \ref{sec:conclusion}.

%Besides being theoretically comprehensive, 
%  the models are also shown, in Section \ref{sec:polsar_models_validation}, %in section \ref{sec:improve_the_match_bw_theory_practice},
%  to be robust in handling practical data. % imperfection that violate conventional assumptions. 
%shows how the statistical models for POLSAR can be considered as the generic
%And since POLSAR can be viewed as a multi-dimensional extension of SAR, the paper

%There are a few benefits of using the homoskedastic measures of distance for POLSAR.
%Firstly, compared to the multiplicative dis-similarity measure of ratio, these additive measures of distance fit more naturally with the linear scale nature of digital imagery.
%Secondly, within the homoskedastic log-transformed domain, where the distance concept as well as the Gauss Markov theorem is again applicable, the universal MSE deserves to be reviewed as a reliable objective function / criteria for (POL)SAR estimators.
%This may open up a number of different research directions for future studies of (POL)SAR.
%As example demonstrations of how these models and distance measures can be used, a novel clustering algorithm, a new speckle filter and various procedures in using MSE to evaluate speckle filters are also presented as part of the thesis.

%Specifically, the rest of this paper is structured as follows.
%After the related work in literature is critically reviewed in Section \ref{sec:lit_review},
%  Section \ref{sec:theoretical_model} presents the theoretical models for both partial and full polarimetric SAR and their derived discrimination measures.
%The representative power of the scalar models are illuminated in Section \ref{sec:sar_special_case_of_polsar},
%  which details how the proposed models for the multi-variate POLSAR also includes the well-known models for the traditional SAR as its special univariate case.
%can be used to derive the well-known models for the univariate SAR data as its special case.
%Besides being theoretically comprehensive, 
%  the model is also shown, in Section \ref{sec:polsar_models_validation}, %in section \ref{sec:improve_the_match_bw_theory_practice},
%  to be robust in handling practical data. % imperfection that violate conventional assumptions. 
%To demonstrate the application of these scalar and representative models, 
%  Section \ref{sec:evaluating_polsar_filters} explores how they can be used to evaluate POLSAR speckle filters.   
%In the end, Section \ref{sec:discussion_conclusion} provides the final conclusion of the paper.
%critically reviews and discusses related literature  
%  before the paper is concluded in Section \ref{sec:conclusion}.

%In this paper, the above results are exploited in the context of POLSAR data processing. In particular, a few scalar, consistent and homoskedastic measures of distance are proposed. The paper is structured as follows:
%  after the preceding introduction to the basics of POLSAR statistical analysis,
%  the second section briefly reviews related work in published literature.
%Section \ref{sec:polsar_heterosked_model_and_log_transform} presents the theoretical models together with conclusive evidence for two related points.
%The first point is that: POLSAR data is multiplicative and heteroskedastic in its original domain.
%And the second conclusion being: log-transformation converts it into an additive and homoskedastic model.
%Consequently a few consistent measures of distance are presented in section \ref{sec:distance_measure}.
%After the model is presented for both partial and full polarimetric SAR data,
%  section \ref{sec:sar_special_case_of_polsar} discusses
%    how the proposed models for the multi-variate POLSAR can be used to derive the well-known models for the univariate SAR data as its special case.
%Besides being theoretically comprehensive, 
%  the model is also shown, in Section \ref{sec:polsar_models_validation}, %in section \ref{sec:improve_the_match_bw_theory_practice},
%  to be robust in handling practical data. % imperfection that violate conventional assumptions. 
%To demonstrate the application of the consistent and homoskedastic dis-similarity measures, 
%  Section \ref{sec:evaluating_polsar_filters} explores how they are used for evaluating POLSAR speckle filters.   
%Finally, Section \ref{sec:discussion_conclusion} critically reviews and discusses related literature  
%  before the paper is concluded in Section \ref{sec:conclusion}.

@@

This section outlines related published work.
In particular, section \ref{sec:lit_models} reviews various scalar statistical models used for different POLSAR observables.
Its purpose is to show that none of those  models have led to statistically consistent discrimination measures. 
Section \ref{sec:lit_measures} further strengthens those findings by discussing discrimination measures that have been proposed for POLSAR,
  and demonstrates that almost all of them are based on the likelihood statistical test for complex Wishart distribution.
However, while an exact statistical distribution is needed for the test,
  only an symptotic distribution is used in the original work \cite{Conradsen_2003_TGRS_4}.

%it should be noted that while the original proposed work \cite{Conradsen_2003_TGRS_4}
%  for the test requires an exact distribution, only an asymptotic distribution is used.  
%This section critically review the related work in published literature.
%Specifically, the first sub-section reviews various scalar statistical models for different POLSAR observables.
%Its purpose is to show that none of the proposed models have led to statistically consistent discrimination measures.
%%shows that while many
%The second sub-section strengthen this findings by reviewing many different discrimination measures that have been proposed for POLSAR data.
%It shows that almost all of them depends on the likelihood statistical test for complex Wishart distribution.
%Unfortunately, in the original proposed work \cite{Conradsen_2003_TGRS_4} for the test,
%  while an exact distribution is to be expected, only an asymptotic distribution was given.

\subsection{Investigated Scalar Observables for POLSAR Data and their Statistical Models}
\label{sec:lit_models}

%Of course, discrimination measures are not the only way to carry out classification tasks in POLSAR.
%Another common approach is to carry out classification using POLSAR observables identified by different target decomposition theorems.
Different target decomposition theorems have identified many possible scalar observables for complex POLSAR data.
In \cite{Alberga_2008_IJRS_4129}, the performance of different scalar POLSAR observables is evaluated for classification purposes.
While many scalar observables for POLSAR were presented, their corresponding statistical models and classifiers were not available.
Furthermore, at its conclusion, the paper indicated that it is impossible to identify a single best representation.
  %``it is impossible to indentify the best one''
  %as these ``representations ... do not show a precise trend''.
%Instead an MLP classification approach were employed. 
Although, to be fair, those observables were identified to describe a decomposed portion of the complex POLSAR data,
  rather than providing a single unified representation of the POLSAR data.
%and thus they are not to be used individually for the purpose of single-handedly representing POLSAR data.

@@

More recently, %other studies has offered
  %\cite{Lopez-Martinez_2003_TGRS_2232} and \cite{Erten_2012_Sensors_2766} gives
  statistical models for
  each element of the POLSAR covariance matrix $S_{pq}S_{rs}^*$ \cite{Lopez-Martinez_2003_TGRS_2232}
  as well as for the largest eigen-value of the covariance matrix $\lambda_1$ \cite{Erten_2012_Sensors_2766} have been studied.
While these models undoubtedly help in understanding POLSAR data,
  none of these observables have been shown to meet the dual criteria of
  1) resulting in statistically consistent discrimination measures and
  2) being representative of the complex POLSAR data.
%While these models undoubtedly help in understanding the SAR data,
%  individually none of these observables has been shown to statisfy the dual criteria of
%  1) resulting in statistically consistent discrimination measures and
%  2) being representative of the complex POLSAR data.
  %that they model have the power to represent the multi-demensional POLSAR covariance matrix.

@@

%This sub-section reviews some relevant and published works relating to different measures of distance for POLSAR data. 
%In particular, a few different matrix distances have been proposed and evaluated in recent review papers\cite{Dabboor_2013_IJRS_1492}\cite{Kersten_2005_TGRS_519}
%will be briefly reviewed.
%specifically a few different matrix distances have been proposed and evaluated in recent review papers\cite{Dabboor_2013_IJRS_1492}\cite{Kersten_2005_TGRS_519}

The commonly used measure of distance for matrices are either the Euclidean or Manhattan distances, defined as:
%The commonly used measure of distance for matrices are either the Euclidean or the Manhattan distance,
%  which are defined in the following equations, respectively:
\begin{align}
  d(C_x,C_y) &= \sum_{i,j} |\mathbb{R} (C_x - C_y)_{i,j}| + \sum_{i,j} |\mathbb{I} (C_x - C_y)_{i,j}| \\
  d(C_x,C_y) &= \sqrt{\sum_{i,j} |C_x - C_y|_{i,j}^2 }
\end{align}
where $C_{i,j}$ denotes the (i,j) elements of the matrix C,
 $||$ denotes absolute values
and $\mathbb{R},\mathbb{I}$ denote the real and imaginary parts respectively.
However, in the context of POLSAR% covariance matrix
, these dis-similarity measures are not widely used 
  probably because of the multiplicative nature of the noisy data.
%In the context of polsar covariance matrix, however, they are not widely used 
%  probably because of the multiplicative nature of the data.

In the field of POLSAR, the Wishart distance is probably the most widely used, as part of the well-known Wishart classifier \cite{Lee_1999_TGRS}.
It is defined \cite{Lee_1994_IJRS_2299} as;
\begin{equation}
  d(C_x,C_y) = \ln|C_y| + tr(C_xC_y^{-1})
\end{equation}
As a measure of distance, its main disadvantage is that $d(C_y,C_y) = \ln|C_y| \neq 0$.

Recent works have suggested alternative dissimilarity measures including the asymmetric and symmetric refined Wishart distance \cite{Anfinsen_2007_ESA_POLINSAR},
\begin{align}
  d(C_x,C_y) &= \frac{1}{2} tr(C_x^{-1}C_y + C_y^{-1}C_x) - d \\
    d(C_x,C_y) &= \ln|C_x| - \ln|C_y| + tr(C_xC_y^{-1}) - d
\end{align}
the Bartlett distance \cite{Kersten_2005_TGRS_519},
  \begin{align}
  d(C_x,C_y) &= 2 \ln |C_{x+y}| - \ln |C_x| - \ln |C_y| - 2d\ln2
  \end{align}
the Bhattacharyya distance \cite{Lee_2011_IGARSS_3740},
\begin{equation}
  r(C_x,C_y) = \frac{|C_x|^{1/2} |C_y|^{1/2}}{|(C_x+C_y)/2|}
\end{equation}
and the Wishart Statistical test distance \cite{Cao_2007_TGRS_3454}
\begin{equation}
  d(C_x,C_y) = (L_x + L_y) \ln|C| - L_x \ln|C_x| - L_y\ln|C_y|
\end{equation}
.

@@

%In comparison to these formulae, the distance measures for two covariance matrices proposed in this paper can be written as:
%\begin{equation}
%  d(C_x,C_y) =  \ln|C_x| - \ln|C_y| 
%\end{equation}
%At first glance, while the propose formula may be a lot simpler, it still operates in a similar way to others in making use of the log-determinant.

Closer investigation of these dis-similarity measures reveals that most are related in some way. %to each other.
The Bhattacharyya distance is easily shown to be related to the Barlett distance.
At the same time the Barlett distance can be considered a special case of the Wishart Statistical Test distance,
  when the two data sets have the same number of looks, i.e. $L_x=L_y$.
%It is also intuitively trivial to arrive at the conclusion that the contrast measure of distance proposed in this paper has fixed statistical behaviour, by comparing the symmetric and asymmetric versions of the refined Wishart distance, assuming that they are shown to follow fixed distributions.
The close relation among the measures is further supported by the fact that
  all of their proposing papers referenced the same statistical model developed in \cite{Conradsen_2003_TGRS_4} as a foundation.
%Interestingly, the change detection model in \cite{Conradsen_2003_TGRS_4} also make use of determinant ratio and log determinant in its derivation.  
%The next sub-section will illustrate how the log-determinant model presented in this paper can offer an alternative and simple derivation for the change-detection statistics.
In \cite{Conradsen_2003_TGRS_4}, to determine if the two scaled multi-look POLSAR covariance matrix $Z_x$ and $Z_y$,
  which have $L_x$ and $L_y$ as the corresponding number of looks,
  come from the same underlying stochastic process,
the likelihood ratio statistics for POLSAR covariance matrix is considered:  
\begin{equation*}
  Q = \frac{(L_x+L_y)^{d \cdot (L_x+L_y)}}{L_x^{d \cdot L_x} L_y^{d \cdot L_y}} \frac{|Z_x|^{L_x} |Z_y|^{L_y} }{|Z_x+Z_y|^{(L_x+L_y)}}
\end{equation*}

Taking the log-transformation of the above statistics, and note that $C_{vx} = Z_x / L_x$, $C_{vy} = Z_y / L_y$ and $C_{vxy} = (Z_x + Z_y)/(L_x + L_y)$ it becomes 
\begin{eqnarray}
  Q &=& \frac{|C_{vx}|^{L_x} \cdot |C_{vy}|^{L_y} }{|C_{vxy}|^{L_x + L_y}} \label{eqn:ori_likelyhood_stats} \\
  \ln Q &=& L_x \ln |C_{vx}| + L_y \ln |C_{vy}| - (L_x + L_y) \ln |C_{vxy}| \label{eqn:log_likelyhood_stats}
\end{eqnarray}

To detect changes, a test statistic is developed based on this measure of distance.
This means a distribution is to be derived for the dissimilarity measure.
However, in the original work \cite{Conradsen_2003_TGRS_4}, only an asymptotic distribution has been proposed.
%In this sub-section, a statistical model is developed for this measure of distance.
%Since both $Z_x$ and $Z_y$ follow a complex Wishart distribution with $L_x$ and $L_y$ degrees of freedom,
%  $Z_x+Z_y$ also follows the complex Wishart distribution with $(L_x + L_y)$ degrees of freedom.
%In view of the models for $|C_v|$ and $\ln|C_v|$ developed in this paper (Eqns. \ref{eqn:determinant_distribution} \& \ref{eqn:log_determinant_distribution}),
%  it is evident that not only the bound for $\ln Q$, or equivalently $Q$, can be derived
%  but the whole statistical distribution for it can be simulated as well.

@@

In this section, several scalar statistical models for POLSAR are presented.
After section \ref{sec:model_basic} briefly introduces the basic foundations, concepts and notation of this paper, %reviews related work in published literature,
  section \ref{sec:polsar_heterosked_model_and_log_transform}
  presents the theoretical models together with mathematical evidence for two related points.
The first point is that the POLSAR data is multiplicative and heteroskedastic in its original domain.
And the second conclusion states that log-transformation converts it into an additive and homoskedastic model.
Some consistent measures of distance are then derived and presented in section \ref{sec:distance_measure}.

@@

%As the covariance matrix is only defined on multiple data points,
%  sample POLSAR covariance matrices are commonly presented in ``ensemble'' format.
The sample POLSAR covariance matrix is formed as the mean of Hermitian outer product of independent single-look scattering vectors,
\begin{equation}
  C_v = \langle ss^{*T} \rangle = \frac{1}{L} \sum^L_{i=1}s_is_i^{*T}
\end{equation}
where $s_i$ denotes the single-look scattering vector,
  which equals $s_{part}$ for the case of partial POLSAR or
  $s_{full}$ for the case of full polarimetry,
and $L$ is the number of looks.

@@

%\section{Handling certain discrepancy between the theoretical model and pratical data}
\subsection{Difference between the theoretical assumptions and the conditions found in practice}  
\label{sec:improve_the_match_bw_theory_practice}
